
---
title: "Overview"
format:
  html:
    theme: flatly
    toc: true
    toc-location: right
    toc-depth: 3
    number-sections: false
    code-fold: false
    code-tools: false
    fig-cap: true
---


```{r quarto setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE)
#| warning: false
#| echo: false
```

```{r Define the Catchment}


# Load our libraries & key datasets

library(sf)
library(tidyverse)
library(DT)

 source("Catch_Set_Up.R")
```

The Wessex Area has been subdivided into 12 ‘Operational Catchments’. Here we focus on the `r unique(CAT$OPCAT_NAME)`, the figures below provide information on its geography and waterbodies, both "unmodified" and Artificial or Heavily Modified Water Bodies (A/HMWB). 


### Surface Water Catchments
The `r unique(CAT$OPCAT_NAME)` Operational Catchment sits within the `r unique(CAT$MNCAT_NAME)` Management Catchment, which sits within the larger `r unique(CAT$RBD_NAME)` River Basin District. It contains `r length(unique(CAT$WB_NAME))` individual EA waterbodies.


```{r OS Style Leaflet Map}
#| warning: false

library(magrittr)
library(ggspatial)
library(leaflet)


#Load in files
  DRN <- read_sf("/dbfs/mnt/lab/unrestricted/harry.gray@environment-agency.gov.uk/DRN/DRN_Merged_MCAT.shp")
  Rivers_250k <- read_sf("/dbfs/mnt/lab/unrestricted/harry.gray@environment-agency.gov.uk/CEP/Hydrology/Rivers_250k.shp")

  DRN <- DRN[CAT,]
# Might need to install 
  
  leaflet() %>% 
    addProviderTiles(providers$OpenStreetMap) %>% 
    addPolygons(data= CAT,
                fillColor = "white",
                color = "black",
                opacity = 1,
                fillOpacity = 0.0000001,
                weight = 1.3,
                popup = ~WB_NAME) %>% 
    addPolylines(data=DRN,
                 color= "steelblue",
                 weight= 1.4,
                 group = "Detailed River Network") %>% 
    addLayersControl(overlayGroups = "Detailed River Network",
                      position = "topright",
                      options= layersControlOptions(collapsed=FALSE)) %>% 
                     hideGroup("Detailed River Network")
       
  
```

```{r Groundwater Catchments}
#import groundwater catchments dataset 
GWC <- read_sf("/dbfs/FileStore/WSX_HGray/ETL_Imports_Require_Manual/WFD_Groundwater_Operational_Catchments_Cycle_3.shp") 

#filter groundwater catchments dataset to just those in the parrett catchment  
WxGWC <- GWC[GWC$OPCAT_ID == "1238" |GWC$OPCAT_ID == "1294" | GWC$OPCAT_ID == "1072" | GWC$OPCAT_ID == "1088" | GWC$OPCAT_ID == "1089",]


#transform coordinates to WGS64 
 WxGWC2 <- st_transform(WxGWC, 4326)
 
  leaflet() %>% 
    addProviderTiles(providers$OpenStreetMap) %>% 
    addPolygons(data= WxGWC,
                fillColor = "white",
                color = "black",
                opacity = 1,
                fillOpacity = 0.0000001,
                weight = 1.3,
                popup = ~OPCAT_NAME) %>% 
    addPolygons(data=CAT_Union)
```

### Geology & Topography of Catchment

The geology of a catchment determines the presence of groundwater aquifers, a catchment's topography, and characteristics of the streams it contains (lowland clay river, chalk streams etc). The characteristics given to a river by its geology (such as natural alkalinity, substrate type etc) impact on biotic and phys-chem indicies used within WFD classifications. 

```{r Topo & geological map}
#| column: screen-inset-shaded
#| layout-nrow: 1
#| fig-width: 11
#| fig-height: 10


library(elevatr)

# CAT %<>% st_transform(st_crs(4326))
 
elev_rast <- elevatr::get_elev_raster(CAT, z=9, clip="locations")

elev_rast_df <- as.data.frame(raster::rasterToPoints(elev_rast))
colnames(elev_rast_df) <- c("x", "y", "Elev")
  
# Here we assign wbs to upper, mid and lower catchment based on topography
#elev_rast_df_3 <- elev_rast_df %>%
#  mutate(TopoRange = cut(Elev,
  #                       breaks = quantile(Elev, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE),
   #                      labels = c("Low", "Medium", "High"),
    #                     include.lowest = TRUE))

#  o <- st_join(CAT, elev_rast_df_3)

# CAT & DRN need the same CRS
CAT <- st_transform(CAT, crs = st_crs(DRN))

# Filter DRN by CAT
DRN <- DRN[CAT,]

ggplot(data = NULL) +
  geom_raster(data = elev_rast_df, aes(x = x, y = y, fill = Elev)) +
 # geom_sf(data = CAT_Union, col = 1, fill = NA) +
  geom_sf(data= DRN, col="#3944bc", fill=NA) +
  scale_fill_gradientn(colours = terrain.colors(200)) +  # 200 to try get the right gradie
  labs(fill = "Elevation (m)") +
  theme_void()+
   annotation_scale(location = "bl", width_hint = 0.5)+
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
                         style = north_arrow_fancy_orienteering)+
  labs(title = paste0(CAT$OPCAT_NAME, " Topographic Map"))




# Convert CAT crs back to WGS84

CAT %<>% st_transform(st_crs(4326))

Geol <- read_sf("/dbfs/mnt/lab/unrestricted/harry.gray@environment-agency.gov.uk/CEP/Geology/WSX_625Geological.shp")


Geol_cols_LEX <- c(
  "LAMBETH GROUP" = "#e6194b",  
  "GAULT FORMATION AND UPPER GREENSAND FORMATION (UNDIFFERENTIATED)" = "#3cb44b",  
  "GREY CHALK SUBGROUP" = "#ffe119",  
  "LIAS GROUP" = "#4363d8", 
  "INFERIOR OOLITE GROUP" = "#f58231",  
  "PENNINE LOWER COAL MEASURES FORMATION AND SOUTH WALES LOWER COAL MEASURES FORMATION (UNDIFFERENTIATED)" = "#911eb4", 
  "DINANTIAN ROCKS (UNDIFFERENTIATED)" = "#42d4f4",  
  "LOWER GREENSAND GROUP" = "#f032e6",
  "GREAT OOLITE GROUP" = "#bfef45",  
  "CORALLIAN GROUP" = "#fabebe",  # pink
  "TRIASSIC ROCKS (UNDIFFERENTIATED)" = "#469990",  # teal
  "WHITE CHALK SUBGROUP" = "#e6beff",  # lavender
  "WARWICKSHIRE GROUP" = "#9A6324",  # brown
  "LOWER DEVONIAN ROCKS (UNDIFFERENTIATED)" = "#fffac8",  # light yellow
  "KELLAWAYS FORMATION AND OXFORD CLAY FORMATION (UNDIFFERENTIATED)" = "#800000",  # maroon
  "BRACKLESHAM GROUP AND BARTON GROUP (UNDIFFERENTIATED)" = "#aaffc3",  # mint
  "UNNAMED EXTRUSIVE ROCKS, SILURIAN" = "#808000",  # olive
  "SOLENT GROUP" = "#ffd8b1",  # peach
  "PURBECK LIMESTONE GROUP" = "#000075",  # navy
  "LLANDOVERY ROCKS (UNDIFFERENTIATED)" = "#a9a9a9",  # dark gray
  "WENLOCK ROCKS (UNDIFFERENTIATED)" = "#d3d3d3",  # light gray
  "PERMIAN ROCKS (UNDIFFERENTIATED)" = "#ff69b4",  # hot pink
  "PORTLAND GROUP" = "#b5651d",  # sienna
  "MILLSTONE GRIT GROUP [SEE ALSO MIGR]" = "#b0e0e6",  # powder blue
  "UPPER DEVONIAN ROCKS (UNDIFFERENTIATED)" = "#f4a460",  # sandy brown
  "MIDDLE DEVONIAN (UNDIFFERENTIATED)" = "#2e8b57",  # sea green
  "THAMES GROUP" = "#8a2be2",  # blue violet
  "SILURIAN ROCKS (UNDIFFERENTIATED)" = "#ff7f50",  # coral
  "HOLSWORTHY GROUP" = "#d2691e",  # chocolate
  "TREMADOC ROCKS (UNDIFFERENTIATED)" = "#ff6347",  # tomato
  "PRIDOLI ROCKS (UNDIFFERENTIATED)" = "#4682b4",  # steel blue
  "WEST WALTON FORMATION, AMPTHILL CLAY FORMATION AND KIMMERIDGE CLAY FORMATION (UNDIFFERENTIATED)" = "#d2b48c",  # tan
  "SOUTH WALES UPPER COAL MEASURES FORMATION" = "#556b2f",  # dark olive green
  "WEALDEN GROUP" = "#8b4513",  # saddle brown
  "LUDLOW ROCKS (UNDIFFERENTIATED)" = "#b8860b",  # dark golden rod
  "PENNINE MIDDLE COAL MEASURES FORMATION AND SOUTH WALES MIDDLE COAL MEASURES FORMATION (UNDIFFERENTIATED)" = "#20b2aa",  # light sea green
  "TEIGN VALLEY GROUP" = "#9932cc"  # dark orchid
)

# Match colours to a custom column

Geol_Cols_RCS_D <- c("CLAY, SILT, SAND AND GRAVEL" = "#b8860b",                           
                     "MUDSTONE, SANDSTONE AND LIMESTONE" = "#f4a460",
                     "CHALK"  = "#aaffc3",
                     "MUDSTONE, SILTSTONE, LIMESTONE AND SANDSTONE" = "#ffd8b1",
                     "LIMESTONE, SANDSTONE, SILTSTONE AND MUDSTONE"  = "#90e6f8",
                     "MUDSTONE, SILTSTONE, SANDSTONE, COAL, IRONSTONE AND FERRICRETE" = "#d2693e",
                     "LIMESTONE WITH SUBORDINATE SANDSTONE AND ARGILLACEOUS ROCKS" = "#d3d3d3",  
                     "SANDSTONE AND MUDSTONE"  = "#ff7f70",
                     "SANDSTONE, LIMESTONE AND ARGILLACEOUS ROCKS" = "#ff7f50",
                   "MUDSTONE, SILTSTONE AND SANDSTONE" = "#ffbf89",
                   "SANDSTONE AND CONGLOMERATE, INTERBEDDED" = "#f86413",
                    "SAND, SILT AND CLAY"   = "#fbd894",                                       
                    "MAFIC LAVA AND MAFIC TUFF"   = "seagreen",                                 
                    "CLAY, SILT AND SAND" = "#b8865e",                                          
                    "LIMESTONE AND MUDSTONE, INTERBEDDED" = "skyblue",                       
                    "LIMESTONE AND CALCAREOUS SANDSTONE"  =   "#c29ea1",                       
                    "LIMESTONE, MUDSTONE AND CALCAREOUS MUDSTONE"  = "#add8e6",               
                    "SANDSTONE AND SILTSTONE, INTERBEDDED" ="#ff9842" )


Geol <- Geol %>%
  mutate(color_rcs = Geol_Cols_RCS_D[RCS_D],
         color_lex = Geol_cols_LEX[LEX])

# Crop the Geology data by catchment
Geol_CAT <- Geol %>% st_intersection(CAT)

#Plot 
 ggplot()+
  geom_sf(data=Geol_CAT, aes(fill=RCS_D))+
  scale_fill_manual( values = setNames(Geol_CAT$color_rcs, Geol_CAT$RCS_D))+
  annotation_scale(location = "bl", width_hint = 0.5)+
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
                         style = north_arrow_fancy_orienteering)+
  theme_void()+
  labs(title = paste0("Bedrock Geology of ", CAT$OPCAT_NAME),
       fill = "Lithology")+
    theme(legend.text = element_text(size = 5.8),
          legend.box.margin = margin(6,6,6),
        legend.position = "bottom")

```


:::{.column-body-outset}
### Land Use
```{r Land Use}
#| fig-width: 17
#| fig-height: 12
#| 
WSX <- read_sf("/dbfs/mnt/lab/unrestricted/harry.gray@environment-agency.gov.uk/Wessex_LC2.shp")

# Filter polygons to catchment

LC <- WSX[CAT,]

# Assign UKCEH colours and habitat classes
  colo=c("#FF0000","#005000","#732600","#00FF00","#01FF7C","#70A800", "#998100", "#FFFF00","#801A80",
         "#E68CA6","#008073", "#D2D2FF","#00005C","#0000E1","#CCB300","#CCB300","#FFFF80","#FFFF80","#8080FF",
         "#000000","#808080")
  
  Hab_Class <- c("Deciduous Woodland", "Coniferous Woodland", "Arable", "Improved Grassland", "Neutral Grassland",
                 "Calcareous Grassland", "Acid Grassland", "Fen", "Heather", "Heather Grassland", "Bog",
                 "Inland Rock", "Saltwater", "Freshwater", "Supralittoral Rock", "Supralittoral Sediment",
                 "Littoral Rock", "Littoral Sediment", "Saltmarsh", "Urban", "Suburban")

# Transform data and assign a palette
   LC$f_mode <- factor(LC$f_mode, levels = 1:21, labels = Hab_Class)
   
   pal_LC <- colorFactor(palette = colo, domain = LC$f_mode)
   
# Create a map of land cover
   
   leaflet() %>%
      addProviderTiles(providers$Esri.WorldImagery) %>%
      addPolygons(data= LC,
                  fillOpacity = 0.7,
                  fillColor = ~pal_LC(f_mode),
                  weight = 0.1,
                  color = NULL,
                  popup = paste0(LC$f_mode," ")) %>% 
      addPolylines(data=DRN,
                   col= "steelblue",
                   weight=3) %>% 
      addScaleBar() %>% 
      addLegend(pal = pal_LC, values = LC$f_mode, opacity = 0.7,
                        title="UKCEH Land Cover")

```
:::

### Bathing Waters Map

The below map shows current riverine and coastal bathing waters. There are no bathing waters in the Parrett.
```{r bathing waters map}
# bathing waters 

BW <- readxl::read_excel("/dbfs/FileStore/WSX_HGray/Bathing_Water_Sites_Comms_Phil_C_Nov_24.xlsx")

cen_coords <- st_coordinates(st_centroid(CAT_Union))

leaflet() %>%
      addProviderTiles(providers$Esri, group="Esri Basemap")%>%
      setView(lng= cen_coords[1], lat = cen_coords[2], zoom=10) %>% 
      addPolygons(data=CAT_Union, 
                  fillOpacity =  0.00001,
                  color = "black") %>% 
      addMarkers(data=BW,
                      popup = ~paste0("Bathing Water: ", BW$SMPT_SHORT_NAME, 
                                      "<br> Priority: ", BW$PRIORITY))  
      
```

### Overall WFD Status
Although the rest of this document provides access to WFD data from 2022, this was an interim WFD year and omitted Chemical classifications and Priority Substances. The most recent full WFD assessment displayed below, occurred 5 years ago in 2019.

Recent classifications can still be rolled-over due to a hiatus in monitoring in certain catchments, this can mean that updated classifications aren't representative of environmental conditions in that year but are an interpretation of previous year's classifications. Figure X [Insert hyperlink here] shows the waterbodies for which classifications have been rolled over [insert below]. 

:::{.column-body-outset}
```{r}

CDE <- read.csv("/dbfs/mnt/lab/unrestricted/harry.gray@environment-agency.gov.uk/CEP/WFD_Wessex_2024.csv")

CDE %<>% 
  filter(Operational.Catchment == unique(CAT$OPCAT_NAME)) %>% 
  inner_join(CAT_geo, ., by = c("WB_ID" = "Water.Body.ID"))

pal <- colorFactor(
  palette = c("green", "seagreen", "seagreen", "yellow", "orange", "#b71105", "red"),
  levels = c("High", "Good", "Supports Good", "Moderate", "Bad", "Poor", "Fail"),
  na.color = "transparent"
)

  ov <- CDE %>% filter(Year == "2019" &
                         Classification.Item ==  "Overall Water Body")

 leaflet() %>% 
  addProviderTiles(providers$Esri,
                  group = "Esri Basemap") %>% 
  addPolygons(data= ov,
              color = "black",
              weight = 0.5,
              fillOpacity = 0.9,
              fillColor = ~pal(Status),
              popup= ~Water.Body) %>% 
   addLegend(opacity = 0.9, 
            pal = pal,
            values = ov$Status,
            title = paste0(unique(ov$Classification.Item),
                           "<br> 2019 Status")) %>%        
  addLayersControl(overlayGroups = c(
             "Esri Basemap"),
                   position = "topright",
                   options= layersControlOptions(collapsed=FALSE)) %>% 
  hideGroup(c("Esri Basemap")) 


```
:::