---
title: "Physical Habitat"
format:
  html:
    theme: flatly
    toc: true
    toc-location: right
    toc-depth: 3
    code-fold: false
    code-tools: false
    fig-cap: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, echo=FALSE, cache=TRUE)
```



```{r}

library(sf)
library(tidyverse)
library(lubridate)
library(magrittr)


source("Catch_Set_Up.R")

 CAT_WBs <- unique(CAT$WB_ID)
 
 DRN <- read_sf("/dbfs/mnt/lab/unrestricted/harry.gray@environment-agency.gov.uk/DRN/DRN_Merged_MCAT.shp")

  DRN <- DRN[CAT,]

```

One of the three pillars in the trinity of ecological health is Physical Habitat. Improvements to water quality and flow within a canal won't be as ecologically effective as improving all three. 

Definition of Physical Habitat- 
_there is a real opportunity to tell the WFD morphology narrative here as many don't have sight of this_


### Morphology related RNAGs

```{r}
# This code chunk could be deleted?

CPS <- read.csv("/dbfs/mnt/lab/unrestricted/harry.gray@environment-agency.gov.uk/CEP/CPS_Exports/WSX_rnags_01082024.csv")

CAT_geo <- subset(CAT, select = c(WB_ID, geometry))

CPS_sf <- right_join(CAT_geo, CPS, by = "WB_ID")

i <- CPS_sf[CPS_sf$SWMI_NAME == "Physical modification",]

Morph_Pres <- i %>% 
        mutate(
          Year = lubridate::year(CREATED_DATE)
        ) %>% 
         filter( OPCAT_NAME == unique(CAT$OPCAT_NAME))



# We could have a datatable
#DT:: datatable(Morph_Pres, class = "cell-border stripe")


# Transform the sf objects to WGS84
Morph_Pres %<>% st_transform(st_crs(4326))
CAT %<>% st_transform(st_crs(4326))



```

Morphology specific RNAGs within `r unique(CAT$OPCAT_NAME)` are as follows: 

```{r Morphology RNAG}


RNAG_Morph <- RFF %>% 
           filter(
           CLASS_ITEM_NAME == "Morphology")# & 
            # OPCAT_NAME == unique(CAT$OPCAT_NAME))

# Calculate the unique count of WB_NAME in RNAG_Morph
unique_count <- length(unique(RNAG_Morph$WB_NAME))

# Conditional logic to check the count and act accordingly
if (unique_count >= 1) {
  leaflet() %>% 
    addProviderTiles(providers$Esri) %>% 
    addPolygons(data=CAT, 
                label = ~WB_NAME, 
                fill="grey",
                color = "black",
                weight = 2,
                fillOpacity = 0.1) %>% 
    addPolygons(data=RNAG_Morph, 
                fillColor = "salmon",
                color = "black",
                label = ~WB_NAME,
                weight = 2,
                fillOpacity = 0.9,
                popup = ~paste0("RNAG created ", Year)) %>% 
    addLegend(labels = c("Classification Item Morphology", "other RNAG"),
              colors = c("salmon","grey"),
              position = "bottomright") %>% 
    addPolylines(data = DRN, 
                 color = "steelblue",
                 opacity = 1,
                 weight = 0.6)
} else {
  print(paste0("The ", unique(CAT$OPCAT_NAME)," doesn't have any morphology RNAGs, the morphology of its waterbodies' therefore 'Support Good' ecological status."))
}
```


```{r}
RHS_WSX <- read_sf("/dbfs/mnt/lab/unrestricted/harry.gray@environment-agency.gov.uk/CEP/RHS_w_MCATs.shp")

RHS <- RHS_WSX[CAT,]

```

### Waterbodies with Physical Modification as their Significant Waterbody Management Issue

The median RHS Habitat Quality Assessment Score in Wessex is `r summary(RHS_WSX$HQA)[3]`, in the `r unique(CAT$OPCAT_NAME)`,  the median RHS HQA is `r summary(RHS$HQA)[3]`. Low scores often indicate channelisation, whilst high scores indicate a more natural physical habitat.


```{r RHS data & Physical Mods}
library(leaflet)
library(lubridate)
#Script which plots River Habitat Survey data.

RHS <-   RHS %>% 
    mutate(
      Date = dmy(SURVEY_D),
      Year = year(Date)
    ) %>% 

group_by(Year) %>% 
  mutate(
    Year_Mean = mean(HQA)
) %>%
  ungroup() %>% 
  mutate(
    Lat = st_coordinates(geometry)[,2],
    Long = st_coordinates(geometry)[,1]
  )

 
pal2 <- colorBin(palette = "viridis", domain=RHS$HQA)


leaflet() %>% 
  addProviderTiles(providers$Esri.WorldImagery,
                   group = "Esri Satellite Basemap") %>% 
    addPolygons(data=CAT, 
              label = CAT$WB_NAME, 
              fill="grey",
              color = "black",
              weight = 2,
              fillOpacity = 0.1) %>% 
  addPolygons(data=Morph_Pres, 
              fillColor = "salmon",
              opacity = 0.8,
              color = "black",
              label = Morph_Pres$WB_NAME,
              weight = 2,
              fillOpacity = 0.9,
              popup = paste0("RNAG created ", Morph_Pres$Year)) %>% 
       # Here's where I added the RHS dat
addPolylines(data = DRN, 
               color = "steelblue",
               opacity = 1,
               weight = 0.6) %>% 
  addCircleMarkers(data = RHS,
                   lng = ~Long, 
                   lat = ~Lat, 
                   col = ~pal2(RHS$HQA),
                   fill = ~pal2(RHS$HQA),
                   fillOpacity = 1,
                   radius = 5,
                   opacity = 1,
                   popup = paste0("Mean HQA Score: ", RHS$HQA, "<br>",
                                  "WFD WB: ", RHS$WB_NAME),
                   group= "RHS: Habitat Quality Assessment"
                   )%>% 
  addLayersControl(baseGroups = c("Esri Satellite Basemap","Blank Basemap"),
                   overlayGroups = "RHS: Habitat Quality Assessment",
                   position = "topright",
                   options= layersControlOptions(collapsed=FALSE)) %>% 
  
  addLegend(pal = pal2, values = RHS$Year_Mean, opacity = 0.7,
                              title= paste0("Yearly mean of Habitat<br> Quality Assessment<br> (Habitat diversity)"),
            group = "RHS: Habitat Quality Assessment") %>% 
  addLegend(labels = c("SWMI Physical Modification", "Other SWMI"),
            colors = c("salmon", "grey"),
            position = "bottomright") %>%
  hideGroup(c("RHS: Habitat Quality Assessment",
              "Esri Satellite Basemap")) 
  


```

### River Obstacles

```{r}

library(jsonlite)
library(httr)

obs_wsx <-  read_sf("/dbfs/FileStore/WSX_HGray/River_obstructions_WSX.shp")

obs_wsx$TYPE <- sub("ARTIFICIAL", "Artificial", obs_wsx$TYPE)
obs_wsx$TYPE <- sub("NATURAL", "Natural", obs_wsx$TYPE)

# .shp already cropped to Wessex area
  
  # SF merge so can filter by opcat.
  obs_wsx <- obs_wsx %>% st_intersection(CAT)
  

  square_green <-
  makeIcon(iconUrl = "https://www.freeiconspng.com/uploads/green-square-1.png",
           iconWidth = 18,
           iconHeight = 18)
  
  square_maroon <- 
    makeIcon(iconUrl = "https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcRc1gE1hAdte8IAuu4piaQSYZA_-nKNuXDEyg&s",
             iconWidth = 18,
             iconHeight = 18)

  square_grey <- 
    makeIcon(iconUrl = "https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcTj6OSkMVt7i1BRqfbWDG0qhrM0SMhF0cXOfg&s",
             iconWidth = 18,
             iconHeight = 18)
    
  
  
  leaflet() %>% 
  addProviderTiles(providers$Esri) %>% 
  addPolylines(data = DRN, 
               color = "steelblue",
               opacity = 1,
               weight = 0.6) %>% 
    addPolygons(data=CAT, 
              label = CAT$WB_NAME, 
              fill="grey",
              color = "black",
              weight = 2,
              fillOpacity = 0.00001) %>% 
  addCircleMarkers(data = RHS,
                   lng = ~Long, 
                   lat = ~Lat, 
                   col = ~pal2(RHS$HQA),
                   fill = ~pal2(RHS$HQA),
                   fillOpacity = 1,
                   radius = 5,
                   opacity = 1,
                   popup = paste0("Mean HQA Score: ", RHS$HQA, "<br>",
                                  "WFD WB: ", RHS$WB_NAME),
                   group= "RHS: Habitat Quality Assessment"
                   ) %>% 
  addMarkers(data = obs_wsx,
             icon= square_green,
               popup = ~paste0("Name: ", obs_wsx$SITE, 
                               "<br>", "Type: ", obs_wsx$TEXTSTRING,
               "<br>Height: ", obs_wsx$OBS_HEIGHT, " cm?",
               "<br>Fish Pass: ", obs_wsx$FISHPASSTY),
             group="Obstacles") %>%
    
  addLayersControl(overlayGroups = c("RHS: Habitat Quality Assessment",
                   "Obstacles"),
                   position = "topright",
                   options= layersControlOptions(collapsed=FALSE)) %>% 
    
  
  addLegend(pal = pal2, values = RHS$Year_Mean, opacity = 0.7,
                              title= paste0("Yearly mean of Habitat<br> Quality Assessment<br> (Habitat diversity)"),
           group = "RHS: Habitat Quality Assessment") %>% 
  hideGroup("RHS: Habitat Quality Assessment") 
 
  

```

