---
title: "Physical Habitat"
format:
  html:
    theme: flatly
    toc: true
    toc-location: right
    toc-depth: 3
    number-sections: true
    code-fold: false
    code-tools: false
    fig-cap: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, echo=FALSE, cache=TRUE)
```

```{css}
.dataTables_wrapper {
  max-height: 350px;
  overflow-y: auto;
}

.dataTables_wrapper table {
font-size: 11px;
}
```

::: callout-warning
## This page is under construction
:::

```{r}

library(sf)
library(tidyverse)
library(lubridate)
library(magrittr)


source("Catch_Set_Up.R")

 CAT_WBs <- unique(CAT$WB_ID)
 
```

One of the three pillars in the trinity of ecological health is Physical Habitat. Improvements to water quality and flow within a canal won't be as ecologically effective as improving all three pillars, including morphology. 

### Morphology related RNAGs

```{r}
# This code chunk could be deleted?

CPS <- read.csv("/dbfs/mnt/lab/unrestricted/harry.gray@environment-agency.gov.uk/CEP/CPS_Exports/WSX_rnags_01082024.csv")

CAT_geo <- subset(CAT, select = c(WB_ID, geometry))

CPS_sf <- right_join(CAT_geo, CPS, by = "WB_ID")

Morph_Pres <- CPS_sf %>% 
        filter(SWMI_NAME == "Physical modification") %>% 
        mutate(
          Year = lubridate::year(CREATED_DATE)
        ) %>% 
         filter( OPCAT_NAME == unique(CAT$OPCAT_NAME))



```

Due to the design of the Water Framework Directive, morphology is very rarely an RNAG, a better representation of where morphological pressure acts within a catchment is shown in the map below, which records where morphology is a Significant Water Management Issue (SWMI), in layman's terms a SWMI is the pressure which we think puts our ability to achieve WFD at most risk.

Morphology specific RNAGs within `r unique(CAT$OPCAT_NAME)` are as follows: 

```{r Morphology RNAG}


RNAG_Morph <- RFF %>% 
           filter(
           CLASS_ITEM_NAME == "Morphology")# & 
            # OPCAT_NAME == unique(CAT$OPCAT_NAME))

# Calculate the unique count of WB_NAME in RNAG_Morph
unique_count <- length(unique(RNAG_Morph$WB_NAME))

# Conditional logic to check the count and act accordingly
if (unique_count >= 1) {
  leaflet() %>% 
    addProviderTiles(providers$Esri) %>% 
    addPolygons(data=CAT, 
                label = ~WB_NAME, 
                fill="grey",
                color = "black",
                weight = 2,
                fillOpacity = 0.1) %>% 
    addPolygons(data=RNAG_Morph, 
                fillColor = "salmon",
                color = "black",
                label = ~WB_NAME,
                weight = 2,
                fillOpacity = 0.9,
                popup = ~paste0("RNAG created ", Year)) %>% 
    addLegend(labels = c("Classification Item Morphology", "other RNAG"),
              colors = c("salmon","grey"),
              position = "bottomright") %>% 
    addPolylines(data = DRN, 
                 color = "steelblue",
                 opacity = 1,
                 weight = 0.6)
} else {
  print(paste0("The ", unique(CAT$OPCAT_NAME)," doesn't have any morphology RNAGs, the morphology of its waterbodies' therefore 'Supports Good'. In this context, 'Supports Good' merely means these waterbodies are not considered to have pristine morphology."))
}
```




```{r}
RHS_WSX <- read_sf("/dbfs/mnt/lab/unrestricted/harry.gray@environment-agency.gov.uk/CEP/RHS_w_MCATs.shp")

RHS <- RHS_WSX[CAT,]

```

### Waterbodies with Physical Modification as their Significant Waterbody Management Issue

The below map shows waterbodies in the `r unique(CAT$OPCAT_NAME)` which have morphology as their SWMI. Reach-level data on morphology can be viewed by clicking the RHS box in the top right.

The median RHS Habitat Quality Assessment Score in Wessex is `r summary(RHS_WSX$HQA)[3]`, in the `r unique(CAT$OPCAT_NAME)`,  the median RHS HQA is `r summary(RHS$HQA)[3]`. Low scores often indicate channelisation, whilst high scores indicate a more natural physical habitat.


```{r RHS data & Physical Mods}
library(leaflet)
library(lubridate)
#Script which plots River Habitat Survey data.

RHS <-   RHS %>% 
    mutate(
      Date = dmy(SURVEY_D),
      Year = year(Date)
    ) %>% 

group_by(Year) %>% 
  mutate(
    Year_Mean = mean(HQA)
) %>%
  ungroup() %>% 
  mutate(
    Lat = st_coordinates(geometry)[,2],
    Long = st_coordinates(geometry)[,1]
  )

 
pal2 <- colorBin(palette = "inferno", domain=RHS$HQA, reverse = T)


leaflet() %>% 
  addProviderTiles(providers$Esri.WorldImagery,
                   group = "Esri Satellite Basemap") %>% 
    addPolygons(data=CAT, 
              label = CAT$WB_NAME, 
              fill="grey",
              color = "black",
              weight = 2,
              fillOpacity = 0.1) %>% 
  addPolygons(data=Morph_Pres, 
              fillColor = "salmon",
              opacity = 0.8,
              color = "black",
              label = Morph_Pres$WB_NAME,
              weight = 2,
              fillOpacity = 0.9,
              popup = paste0("RNAG created ", Morph_Pres$Year)) %>% 
       # Here's where I added the RHS dat
  addPolylines(data = DRN, 
               color = "steelblue",
               opacity = 1,
               weight = 0.6) %>% 
  addCircleMarkers(data = RHS,
                   lng = ~Long, 
                   lat = ~Lat, 
                   col = "black",
                   weight = 1,
                   fill = ~pal2(RHS$HQA),
                   fillOpacity = 1,
                   radius = 5,
                   opacity = 1,
                   popup = paste0("Mean HQA Score: ", RHS$HQA, "<br>",
                                  "WFD WB: ", RHS$WB_NAME),
                   group= "RHS: Habitat Quality Assessment"
                   )%>% 
  addLayersControl(baseGroups = c("Esri Satellite Basemap","Blank Basemap"),
                   overlayGroups = "RHS: Habitat Quality Assessment",
                   position = "topright",
                   options= layersControlOptions(collapsed=FALSE)) %>% 
  
  addLegend(pal = pal2, values = RHS$Year_Mean, opacity = 0.7,
                              title= paste0("Yearly mean of Habitat<br> Quality Assessment<br> (Habitat diversity)"),
            group = "RHS: Habitat Quality Assessment") %>% 
  addLegend(labels = c("SWMI Physical Modification", "Other SWMI"),
            colors = c("salmon", "grey"),
            position = "bottomright") %>%
  hideGroup(c("RHS: Habitat Quality Assessment",
              "Esri Satellite Basemap")) 
  


```

### Reasons for Not Acheiving Good

The below table shows RNAGs which either have morphology as their primary reason for failure (very rare) or either of their main pressures as morphology (more common).

```{r RNAG Datatable PhysHab}



rnag1 <- RFF %>%  filter(PRESSURE_1 == "Morphology" |  PRESSURE_2 == "Morphology")

rnag2 <- RFF %>%  filter(CLASS_ITEM_NAME == "Morphology") %>% 
                  arrange(WATERBODY_ID)           # Arrange dataframe so it appears neater in DT


RNAG <- rbind(rnag1, rnag2)

RNAG <- RNAG %>% rename(YEAR_RNAG_CREATED= CREATED_DATE,
                        WFD_YEAR = YEAR) %>% 
            distinct()

# We could have a datatable
#DT:: datatable(Morph_Pres, class = "cell-border stripe")


# Transform the sf objects to WGS84
Morph_Pres %<>% st_transform(st_crs(4326))
CAT %<>% st_transform(st_crs(4326))

if(nrow(RNAG) >0 ){
DT::datatable(RNAG[,c(7, 6,18, 32, 16:17,19,20,22:30,23, 10,11,2)],
          class = 'cell-border stripe',
          extensions = c("Buttons","FixedColumns"),
          options = list(
            dom = 'Blfrtip',
            buttons = 'excel',
            initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
    "}"))
    )
}
```

### Measures

```{r Measures}

if(nrow(RNAG)>0){

Mes %<>%  filter(WATERBODY_ID %in% unique(RNAG$WATERBODY_ID))

CAT_Mini <- CAT %>% select(WB_NAME, WB_ID) %>% 
  st_drop_geometry()

Mes_Categ <- inner_join(Mes, Measures_Cat, by = "ACTION_UID") %>% 
             inner_join(CAT_Mini, by = c("WATERBODY_ID" = "WB_ID")) %>% 
              filter(!CREATED_DATE== "01/01/1980" & DEPRECATED_CATEGORY == "No") %>% 
              distinct() %>% 
              arrange(WATERBODY_ID)


   DT::datatable(Mes_Categ[,c(16,3,10,12:14)],
          class = 'cell-border stripe',
          extensions = c("Buttons","FixedColumns"),
          options = list(
            dom = 'Blfrtip',
            buttons = 'excel',
            initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
    "}"))
    )
}
```

### River Obstacles
The below map displays all types of barriers and when clicked on, reveals whether fish/eel passes have been installed.

```{r}

library(jsonlite)
library(httr)

obs_wsx <-  read_sf("/dbfs/FileStore/WSX_HGray/River_obstructions_WSX.shp")

obs_wsx$TYPE <- sub("ARTIFICIAL", "Artificial", obs_wsx$TYPE)
obs_wsx$TYPE <- sub("NATURAL", "Natural", obs_wsx$TYPE)

# .shp already cropped to Wessex area
  
  # SF merge so can filter by opcat.
  obs_wsx <- obs_wsx %>% st_intersection(CAT)
  

  square_green <-
  makeIcon(iconUrl = "https://www.freeiconspng.com/uploads/green-square-1.png",
           iconWidth = 18,
           iconHeight = 18)
  
  square_maroon <- 
    makeIcon(iconUrl = "https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcRc1gE1hAdte8IAuu4piaQSYZA_-nKNuXDEyg&s",
             iconWidth = 18,
             iconHeight = 18)

  square_grey <- 
    makeIcon(iconUrl = "https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcTj6OSkMVt7i1BRqfbWDG0qhrM0SMhF0cXOfg&s",
             iconWidth = 18,
             iconHeight = 18)
    
  
  
  leaflet() %>% 
  addProviderTiles(providers$Esri) %>% 
  addPolylines(data = DRN, 
               color = "steelblue",
               opacity = 1,
               weight = 0.6) %>% 
    addPolygons(data=CAT, 
              label = CAT$WB_NAME, 
              fill="grey",
              color = "black",
              weight = 2,
              fillOpacity = 0.00001) %>% 
  addCircleMarkers(data = RHS,
                   lng = ~Long, 
                   lat = ~Lat, 
                   col = ~pal2(RHS$HQA),
                   fill = ~pal2(RHS$HQA),
                   fillOpacity = 1,
                   radius = 5,
                   opacity = 1,
                   popup = paste0("Mean HQA Score: ", RHS$HQA, "<br>",
                                  "WFD WB: ", RHS$WB_NAME),
                   group= "RHS: Habitat Quality Assessment"
                   ) %>% 
  addMarkers(data = obs_wsx,
             icon= square_green,
               popup = ~paste0("Name: ", obs_wsx$SITE, 
                               "<br>", "Type: ", obs_wsx$TEXTSTRING,
               "<br>Fish Pass: ", obs_wsx$FISHPASSTY),
             group="Obstacles") %>%
    
  addLayersControl(overlayGroups = c("RHS: Habitat Quality Assessment",
                   "Obstacles"),
                   position = "topright",
                   options= layersControlOptions(collapsed=FALSE)) %>% 
    
  
  addLegend(pal = pal2, values = RHS$Year_Mean, opacity = 0.7,
                              title= paste0("Yearly mean of Habitat<br> Quality Assessment<br> (Habitat diversity)"),
           group = "RHS: Habitat Quality Assessment") %>% 
  hideGroup("RHS: Habitat Quality Assessment") 
 
  

```

