---
title: "Water Resources"
format:
  html:
    theme: flatly
    toc: true
    toc-location: right
    toc-depth: 3
    code-fold: false
    code-tools: false
    fig-cap: true
    callout-icon: false
---

::: {.callout-warning}
## This page is under construction
:::

```{r quarto setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE)
#| warning: false
#| echo: false
```

```{css}
.dataTables_wrapper {
  max-height: 350px;
  overflow-y: auto;
}

.dataTables_wrapper table {
font-size: 11px;
}
```

```{r }
#| label: the general cps & cat

library(sf)
library(magrittr)
library(tidyverse)
library(leaflet)
library(lubridate)

source("Catch_Set_Up.R")

```


```{r}


# Load your data
CDE <- read.csv("/dbfs/mnt/lab/unrestricted/harry.gray@environment-agency.gov.uk/CEP/WFD_Wessex_2024.csv")

CDE %<>% 
  filter(Operational.Catchment == unique(CAT$OPCAT_NAME)) %>% 
  inner_join(CAT_geo, ., by = c("WB_ID" = "Water.Body.ID"))

# Define palette
pal <- colorFactor(
  palette = c("green", "seagreen", "seagreen", "yellow", "#c65102", "#b71105", "#b71105","red"),
  levels = c("High", "Good", "Supports Good", "Moderate", "Bad", "Poor", "Does Not Support Good", "Fail"),
  na.color = "transparent"
)

# 2022 map
CDE_e_2022 <- CDE %>% 
  filter(Classification.Item == "Hydrological Regime" & 
           Year == "2022")


```

### Hydrological Regime WFD

Within the `r unique(CAT$OPCAT_NAME)`'s `r length(unique(CAT$WB_NAME))` individual EA waterbodies.

In the most recent WFD classification in `r sort(unique(CDE$Year), decreasing=TRUE)[1]`,  `r round(sum(CDE_e_2022$Status == "Supports Good")/length(unique(CAT$WB_NAME))*100, digits=0)`% of waterbodies scored " Supports Good" for `r unique(CDE_e_2022$Classification.Item)`. Whilst `r round(sum(CDE_e_2022$Status == "High")/length(unique(CAT$WB_NAME))*100, digits=0)`% of waterbodies scored "High".


```{r WFD Map morphology}


CDE_m <- CDE %>% 
  filter(Classification.Item == "Hydrological Regime")

CDE_m_22 <- CDE_m %>% 
         filter(Year == sort(unique(CDE$Year), decreasing=TRUE)[1])

CDE_m_19 <- CDE_m %>% 
         filter(Year == sort(unique(CDE$Year), decreasing=TRUE)[2])

CDE_m_15 <- CDE_m %>% 
         filter(Year == sort(unique(CDE$Year), decreasing=TRUE)[3])

CDE_m_last3 <- CDE_m %>% 
          filter(Year == sort(unique(CDE$Year), decreasing=TRUE)[1] &
                          sort(unique(CDE$Year), decreasing=TRUE)[2] &
                          sort(unique(CDE$Year), decreasing=TRUE)[3]
                 )

 leaflet() %>% 
  addProviderTiles(providers$Esri,
                   group = "Esri Basemap") %>% 
   addPolygons(data= CDE_m_22,
              color = "black",
              weight = 0.5,
              fillOpacity = 0.9,
              fillColor = ~pal(Status),
              popup= ~Water.Body,
              group= "2022 Status") %>% 
  addPolygons(data= CDE_m_19,
              color = "black",
              weight = 0.5,
              fillOpacity = 0.9,
              fillColor = ~pal(Status),
              popup= ~Water.Body,
              group= "2019 Status") %>% 
  addPolygons(data= CDE_m_15,
              color = "black",
              weight = 0.5,
              fillOpacity = 0.9,
              fillColor = ~pal(Status),
              popup= ~Water.Body,
              group = "2015 Status") %>% 
  addLegend(opacity = 0.9, 
            pal = pal,
            values = CDE_m$Status,
            title = unique(CDE_m_last3$Classification.Item)) %>% 
  addLayersControl(baseGroups = c( "2022 Status",
              "2019 Status",
              "2015 Status"),
                   overlayGroups = c(
             "Esri Basemap"),
                   position = "topright",
                   options= layersControlOptions(collapsed=FALSE)) %>% 
  hideGroup(c("2019 Status",
              "2015 Status",
              "Esri Basemap"))


```

### Reasons for Not Acheiving Good

Use the table below to search for a term and scroll through RNAGs.
_Measures are yet to be added to this doc_

```{r RNAGS datatable}

# These are just for ecological quality elements

RNAG <- CPS_sf %>% 
                mutate(
                Date = as.Date(ymd_hms(RNAG_CREATED_DATE))
                ) %>% 
               filter(CLASS_ITEM_NAME =="Hydrological Regime") %>% 
               # filter(hydromorph == "not designated artificial or heavily modified") %>%  # I think we should still consider A/HMWBs
                group_by(PRESSURE_NAME_1) %>% 
                filter(Date == sort(Date)[length(sort(Date))]) %>%   # We want the most recent creation date for each pressure
                ungroup() 


library(DT)
library(htmlwidgets)

DT::datatable(RNAG[,c(1, 11:32)],
          class = 'cell-border stripe',
          caption = 'Use the map below to find a waterbody and use the table search function to find its records',
          options = list(dom = 'Blfrtip', buttons = 'copy')
              ) 
            

```


### RNAGs map


RNAGs for Hydrological Regime Elements are displayed in the map below, they are filtered to show the most recent RNAGs, those which are opaque are >5 years old and have a creation date of 1st January 2019. This surpasses the national KPI as it is four months ahead of the KPI target of 1st April 2019.



```{r Hydrol Regime RNAG Map}


unique_count <- length(unique(RNAG$WB_NAME))

if (unique_count >= 1) {
       leaflet() %>% 
  addProviderTiles(providers$Esri) %>% 
    addPolygons(data=CAT_Union,
                        color="black",
                       weight = 2,
                       opacity = 0.99,
                       fillColor = NA,
                       fillOpacity = 0.00001
                  ) %>% 
    addPolygons(data=RNAG,
              color = "black",
              weight = 0.5,
              fillOpacity = ~ifelse(RNAG$Date < as.Date("2019-01-01"), 0.2, 0.99),
              fillColor = ~pal(RNAG$CLASSIFICATION_NAME),
              popup= paste0("RNAG Created: ",RNAG$Date, "<br>",
                            "Element: ", RNAG$CLASS_ITEM_NAME, "<br>",
                            "WB: ",RNAG$WB_NAME, "<br> Pressure 1: ",
                            RNAG$PRESSURE_NAME_1,
                            "<br> Pressure 2: ",
                            RNAG$PRESSURE_NAME_2,
                            "<br> RFF Action Name: " ,
                            RNAG$RFF_ACTION_NAME),
              group = "Invertebrates RNAGs")%>% 
                addLegend(opacity = 0.9, 
                         pal = pal,
                         values = RNAG$CLASSIFICATION_NAME,
                         title = "RNAG by Phys-Chem Quality Element") 
} else {
  print(paste0("The ", unique(CAT$OPCAT_NAME)," doesn't have any Hydrological Regime RNAGs, the hydrology of its waterbodies' therefore 'Support Good' ecological status."))
}

```

# Add extra sections

_These sections are to convey some granularity_

_input from Chris/Bryony perhaps Ben H_

_Add Groundwater_ 