---
title: "Physical_Hab_July"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Physical Habitat

```{r, echo=FALSE, warnings=FALSE}

library(sf)
library(tidyverse)
library(lubridate)
library(magrittr)


 catch <- read_sf("/dbfs/mnt/lab/unrestricted/harry.gray@enivronment-agency.gov.uk/Interim_WFD_2022.shp")# Catchment shapefiles
 CAT <- catch[catch$OPCAT_NAME == "Brue and Axe",]
 
 CAT %<>% st_transform(4326)
  
 CAT_WBs <- unique(CAT$WB_ID)

```

The third pillar in the trinity of ecological health is Physical Habitat, also known as "the silent killer". 

For morphology scores of a catchment see WFD morphology graph below

# Morphology related RNAGs

Insert a DT here so users can interact with RNAG table results

# WFD Morphology Status 
```{r, echo=FALSE, warnings=FALSE}

#library(odbc)
#library(tidyverse)

#con <- dbConnect(odbc(),
 #               Driver = "SQL Server",
  #              Server = "PRDCPZDBSSQ1402.PRODDS.NTNL",
   #              Database = "CPS_REPORTING_LIVE",
    #             UID = "CPSProdReport",
     #            PWD = "Pr0d{psR3p0r+!ng",
      #           Port = 1433)

#rnag <- dbGetQuery(con, "SELECT CLASSIFICATION_NAME, CLASS_ITEM_NAME, PRESSURE_NAME_1, PRESSURE_NAME_2,
 #                       PROBLEM_VALID_STATUS_NAME, SWMI_NAME, SWMI_CERTAINTY_NAME, RFF_ACTION_NAME, 
  #                      ACTION_CERTAINTY_NAME, BUSINESS_CATEGORY_NAME, BUSINESS_CATEGORY_CERTAINTY_NAME,
   ##                     BUSINESS_SECTOR_NAME, TYPE_1_OUTCOME_NAME, TYPE_2_OUTCOME_NAME, TYPE_3_OUTCOME_NAME,                                                                         SOURCE_APPORTIONMENT_NAME, SOURCE_APPORTIONMENT_TYPE_NAME, EARLIEST_CLASSIFICATION_YEAR, 
     ##                   PERIOD_NAME, WATERBODY_ID, CREATED_DATE
                        
       #                 FROM reporting.vw_WaterbodyRff")  #We call in the database in the dbConnect line. We then 
        #                                                          # call in the 

#RNAG_PHY <- rnag %>% filter(SWMI_NAME== "Physical modification" & 
 #                             PERIOD_NAME == "Cycle 3" &
  #                            WATERBODY_ID == CAT_WBs)
#


```


```{r, echo=F, warnings= F}
RHS_WSX <- read_sf("/dbfs/mnt/lab/unrestricted/harry.gray@enivronment-agency.gov.uk/CEP/RHS_w_MCATs.shp")

RHS <- RHS_WSX[CAT,]


```

# River Habitat Survey Data

The median RHS Habitat Quality Assessment Score in Wessex is `r summary(RHS_WSX$HQA)[3]`, in the `r unique(CAT$OPCAT_NAME)`,  the median RHS HQA is `r summary(RHS$HQA)[3]`, this is `r paste0(((round(summary(RHS$HQA)[3]/summary(RHS_WSX$HQA)[3])-1)*100),"%")` less than the Wessex area median. Low scores often indicate channelisation, whilst high scores indicate ...


For access to more granular data, see individual River Habitat Survey data below merged with EA obstacle data.
Re obstacles However, Jones et al (2019) estimate that EA dataset underestimates stream fragmentation by >68%. To make up this shortfall, Citizen Science RiverObstacles data is also displayed (magenta coloured).



```{r, echo=FALSE, warnings=FALSE}
library(leaflet)
#Script which plots River Habitat Survey data.


DRN <- read_sf("/dbfs/mnt/lab/unrestricted/harry.gray@enivronment-agency.gov.uk/DRN/DRN_Merged_MCAT.shp")

  DRN <- DRN[CAT,]
#Filter for catc
 

RHS <-   RHS %>% 
    mutate(
      Date = dmy(SURVEY_D),
      Year = year(Date)
    ) %>% 

group_by(Year) %>% 
  mutate(
    Year_Mean = mean(HQA)
) %>%
  ungroup() %>% 
  mutate(
    Lat = st_coordinates(geometry)[,2],
    Long = st_coordinates(geometry)[,1]
  )

 
pal2 <- colorBin(palette = "viridis", domain=RHS$HQA)

leaflet() %>% 
  addProviderTiles(providers$Esri) %>% 
  addPolylines(data = DRN, 
               color = "steelblue",
               opacity = 1,
               weight = 0.6) %>% 
  addCircleMarkers(data = RHS,
                   lng = ~Long, 
                   lat = ~Lat, 
                   col = ~pal2(RHS$HQA),
                   fill = ~pal2(RHS$HQA),
                   fillOpacity = 1,
                   radius = 5,
                   opacity = 1,
                   popup = paste0("Mean HQA Score: ", RHS$HQA, "<br>",
                                  "WFD WB: ", RHS$WB_NAME)
                   ) %>% 
  addLegend(pal = pal2, values = RHS$Year_Mean, opacity = 0.7,
                              title= paste0("Yearly mean of Habitat<br> Quality Assessment<br> (Habitat diversity)"))#

### Open Obstacles Data from Riverobstacles, updated every 24hrs.

#The River Obstacles initiative is a joint endeavour by the Environment Agency, Zoological Society of London, The Rivers Trust, Thames Estuary Partnership, The River Restoration Centre and Natural Apptitude.

library(jsonlite)
library(httr)

url <- "https://services3.arcgis.com/Bb8lfThdhugyc4G3/arcgis/rest/services/River_Obstacles_Public_View/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

api <- GET(url)

content <- content(api, "text", encoding = "UTF-8")
  data <- fromJSON(content, flatten = TRUE)
  
  #Create lat/long columns from sf object
  obs <- st_as_sf(data$features, coords = c("geometry.x", "geometry.y"), crs = 4326) 
  
  #Filter to unatural structures and then extrct lat longs
  obs %<>%
    filter(attributes.origin == "man_made") %>%
    mutate(Lat = st_coordinates(.)[, 2],
           Long = st_coordinates(.)[, 1])
  
  # Crop so just Wessex
  obstacs_wsx <- obs[CAT,]
  
  # SF merge so can filter by opcat.
  obs_wsx <- obstacs_wsx %>% st_intersection(CAT)
  
```

